Whole-body connectome of a segmented annelid larva

### 2024-09-07 Manuscript.qmd
MS revision: add refs

The *Platynereis* larval nervous system is subdivided into an anterior brain and a rope-ladder-like paired ventral nerve cord (VNC).
The brain and VNC are connected by circumesophageal connectives (Figure 1C, F-H).
The trunk has five segments, an anterior cryptic segment (segment 0)[@steinmetz2011segmental], three main segments (segments 1-3) with parapodia, and a posterior pygidium (Figure 1A, B).

@@

The three-day-old larva has three main trunk segments with chaeta-bearing parapodia (chaetigerous segments) and a more anterior cryptic segment [@steinmetz2011segmental].
In addition, the pygidium forms the posterior-most part of the body [@starunov2015metameric] (Figure 1A, Figure 11A).
The ciliary bands mark the posterior segment boundaries.

@@

By mapping cell-type distributions across segments we found evidence for segmental homology overlain by a heteromeric pattern with segment-specific cell types.
The serial homology of the cryptic segment (sg0) with the other three trunk segments (sg1-3) [@steinmetz2011segmental] is confirmed by neuron and muscle types shared between these four segments.
In addition, the pygidium and the head may have metameric origin.
The pygidium has a coelomic cavity and muscle cell types (MUSlong, MUStrans) shared with the other trunk segments [@starunov2015metameric; @jasek2022].
The segmental origin of the annelid head is suggested by a Cambrian annelid fossil where the head is formed by a segment with parapodia and chaetae [@parry2015cambrian].

>>>>

The *Platynereis* larval nervous system is subdivided into an anterior brain and a rope-ladder-like paired ventral nerve cord (VNC).
The brain and VNC are connected by circumesophageal connectives (Figure 1C, F-H).
The trunk has five segments, an anterior cryptic segment (segment 0)[@saudemont2008complementary; @steinmetz2011segmental], three main segments (segments 1-3) with parapodia, and a posterior pygidium (Figure 1A, B).

@@

The three-day-old larva has three main trunk segments with chaeta-bearing parapodia (chaetigerous segments) and a more anterior cryptic segment [@saudemont2008complementary; @steinmetz2011segmental].
In addition, the pygidium forms the posterior-most part of the body [@starunov2015metameric] (Figure 1A, Figure 11A).
The ciliary bands mark the posterior segment boundaries.

@@

By mapping cell-type distributions across segments we found evidence for segmental homology overlain by a heteromeric pattern with segment-specific cell types.
The serial homology of the cryptic segment (sg0) with the other three trunk segments (sg1-3) [@saudemont2008complementary; @steinmetz2011segmental] is confirmed by neuron and muscle types shared between these four segments.
In addition, the pygidium and the head may have metameric origin.
The pygidium has a coelomic cavity and muscle cell types (MUSlong, MUStrans) shared with the other trunk segments [@starunov2015metameric; @jasek2022].
The segmental origin of the annelid head is suggested by a Cambrian annelid fossil where the head is formed by a segment with parapodia and chaetae [@parry2015cambrian].
